---
title: "Querying the anatomical and clinical annotation hub"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Querying the anatomical and clinical annotation hub}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

PhysioAnnotationHub is a lightweight, base-R-only annotation hub for the
PhysioExperiment ecosystem. It bundles curated anatomical ontology data
(muscles, bones, nerves), clinical codes (ICD-10, ICF), and an anatomical
knowledge graph as CSV files under `inst/extdata`. Everything here runs offline
-- nothing is downloaded at run time.

```{r setup}
library(PhysioAnnotationHub)
```

## Load the hub

`loadAnnotationHub()` reads all bundled datasets into one cached object; printing
it summarises what is available.

```{r load}
hub <- loadAnnotationHub()
hub
```

## Anatomical queries

Look up a muscle, bone, or nerve by name (matching is case-insensitive and
fuzzy); call a getter with no arguments to get the whole table.

```{r anatomy}
getMuscleAnnotation("Trapezius")[, c("muscle_name", "body_region",
                                     "action_primary", "nerve")]
getNerveAnnotation("Accessory")
```

Clinical codes that reference a muscle:

```{r clinical}
getClinicalCodes("Rectus abdominis", system = "icd10")[, c("icd10_code",
                                                            "description")]
```

## The knowledge graph

Anatomical relationships are stored as subject-predicate-object triples. Query
them by any combination of the three slots:

```{r kg-query}
head(queryKG(subject = "Trapezius"))
```

Traverse the graph -- neighbours within N hops, or the shortest path between two
entities:

```{r kg-traverse}
nrow(kgNeighbors("Trapezius"))
kgShortestPath("Trapezius", "External Occipital Protuberance")
```

Over-representation analysis asks whether a muscle set is enriched for an
annotation (here, body region) relative to the full hub:

```{r kg-enrich}
kgEnrichment(
  c("Trapezius", "Latissimus Dorsi", "Serratus Posterior Superior"),
  annotation_type = "body_region"
)
```

## ICF classification

Map a physiological metric or clinical instrument to WHO ICF categories, or pull
a published ICF Core Set:

```{r icf}
tagICF("gait_speed")
icfCategories(c("b730", "d450"))
```

## Where to go next

* `?loadAnnotationHub` -- the hub object and its tables.
* `?queryKG`, `?kgShortestPath`, `?kgEnrichment` -- graph queries.
* `?tagICF`, `?getCoreSet` -- ICF linking.

This hub is designed to be imported by other ecosystem packages (e.g.
PhysioMSKNet) that need anatomical metadata without carrying their own copy.
